Extreme precip and heat waves: empirical-statistical downscaling and analysis

### 2016-09-19 ClimaTrans/climatrans.Rmd
Added run-off and dipper examples

@@

```{r}
wq95 <- function(x) {x <- x[is.finite(x)]; x <- x[x >= 1]; wq95 <- quantile(x,probs=0.95); wq95}
season<-5:10
fw <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wetfreq')
mu <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wetmean')
nd <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'nv')
ndhr <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'count',threshold=100)
q95 <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wq95')
```

@@

```{r, fig.width=6, fig.height=6}
plot(ndhr,new=FALSE)
grid()
```

@@

```{r fig.width=6, fig.height=6}
ndhr.cal <- data.frame(y = coredata(subset(subset(ndhr,is=2),it=c(1949,2004))),
                       X = coredata(subset(eof.es,it=c(1949,2004))))
ds.ndhr <- glm(y ~ X.1 + X.2 + X.3 + X.4 + X.5 + X.6 + X.7, data=ndhr.cal,family="poisson")
plot(1949:2004,ndhr.cal$y,pch=19,ylab=expression(N(X > 100*mm)),xlab='',main=loc(ndhr)[2])
grid()
lines(1949:2004,exp(predict(ds.ndhr)),col='red')
```

@@

>>>>

### Function to evaluate spatio-temporal variance structure for SLP

```{r}
showceofvar <-  function(x,N,i,gcmnm.i,add=FALSE,xlim=NULL,ylim=NULL,verbose=FALSE,...) {
  n <- length(x$mean.diff)
  j <- 1:n
  col <- rgb(j/n,abs(sin(pi*j/n)),(1-j/n),0.3)
  if (is.null(xlim)) xlim <- c(-1,1)
  if (is.null(ylim)) ylim <- c(1,N)

  if (!add) {
    par(bty="n")
    par0 <- par()
    plot(xlim,ylim,type="n",
         ylab="GCM",xlab=expression(1- sigma[p*r*e]/sigma[r*e*f]),
         main=paste("Diagnostics: common EOFs",attr(x,'variable')),
         xlim=xlim,ylim=ylim,col="grey")
    grid()
    #legend(xlim[1],ylim[2],c("same sign","different sign"),
    #       pch=c(19,21),bty="n",col="grey")
    par(xpd=TRUE)
    text(xlim[2],ylim[2],'EOF #',col='grey40',cex=0.8,pos=3)

    par(new=TRUE,fig=c(0.85,0.95,0.70,0.85),mar=c(0,3,0,0),
        cex.axis=0.7,yaxt="s",xaxt="n",las=1)
     colbar <- rbind(1:n,1:n)
    image(1:2,1:n,colbar,col=col)
    par(fig=par0$fig,mar=par0$mar,cex.axis=par0$cex.axis,
        yaxt=par0$yaxt,xaxt=par0$xaxt,las=par0$las,new=TRUE)
    plot(xlim,ylim,type="n",xlim=xlim,ylim=ylim,
         xlab='',ylab='',main='',sub='')
    par(par0$new)
  }
  cex <- abs(0.1 - x$mean.diff)/0.1
  pch <- rep(19,n); pch[cex < 0] <- 21
  cex <- abs(cex); cex[cex > 2] <- 2
  if (verbose) {
     print('Mean difference:');print(x$mean.diff)
     print('Ration of standard deviation');print(x$sd.ratio)
     print('Size');print(cex)
     print('col');print(col)
     #points(x$mean.diff,1-x$sd.ratio,pch=pch,col='grey75',cex=1)
  }

  points(1-x$sd.ratio,rep(i,n),pch=pch,col=col,cex=cex)
  text(xlim[1],i,gcmnm.i,pos=4,cex=0.7)
}

biasdiag <- function(predictor,param='slp',
                    path="CMIP5.monthly/",pattern="slp_Amon_ens_",type='ncdf',
                    period=c(1950,2015),rcp='rcp45',verbose=TRUE,
                    lon=c(0,100),lat=c(65,90),select=NULL,xlim=NULL) {

  it <- month.abb[as.numeric(rownames(table(month(predictor))))]
  if (verbose) print(paste('biasdiag, it=',it))

  d.y <- 6 # Number of leading EOFs to evalue
  years <- 1900:2100
  m <- length(years)
  months <- rep(month(y)[1],m)
  X <- matrix(rep(NA,N*m*d.y),N,m*d.y)
  dim(X) <- c(d.y[2],N,m)
  gcmnm <- rep("",N)
  scorestats <- matrix(rep(NA,N*9),N,9)
  colnames(scorestats) <- c("1-r.xval","mean.diff","sd.ratio","autocorr.ratio",
                            "res.trend","res.K-S","res.ar1",'amplitude.ration',
                            "1-R2")

  t <- as.Date(paste(years,months,'01',sep='-'))

  cols <- rgb(seq(1,0,length=100),rep(0,100),seq(0,1,length=100),0.15)

  flog <- file("DSensemble.ceof-log.txt","at")

  ## Set up a list environment to keep all the results
  results <- list() 
  if (verbose) print("loop...") 
  for (i in 1:N) {
    if (verbose) print(ncfiles[select[i]])
    gcm <- retrieve(ncfile = ncfiles[select[i]],type=type,
                          lon=range(lon(predictor)),
                          lat=range(lat(predictor)),verbose=verbose)
    if (!is.null(it)) {
      if (verbose) print('Extract some months ot a time period')
      if (verbose) print(it)
      gcm <- subset(gcm,it=it)
    }
    #gcmnm[i] <- attr(gcm,'model_id')
    gcmnm.i <- paste(attr(gcm,'model_id'),attr(gcm,'realization'),sep="-r")
    gcmnm.i <- gsub(' ','',gcmnm.i)
    gcmnm.i <- gsub('-','.',gcmnm.i)
    print(gcmnm.i)
    GCM <- aggregate(subset(gcm,it=month.abb[month(y)[1]]),year,FUN='mean') 
    SLPGCM <- combine(predictor,GCM)
    Z <- try(EOF(SLPGCM))
    diag <- diagnose(Z)
    showceofvar(diag,N,i,gcmnm.i,add=(i!=1),xlim=xlim,verbose=verbose)
    eval(parse(text=paste('results$',it,'.',gcmnm.i,' <- diag',sep='')))
  }
  return(results)
}

@@

```{r}
wq95 <- function(x) {x <- x[is.finite(x)]; x <- x[x >= 1]; wq95 <- quantile(x,probs=0.95); wq95}
season<-5:10
fw <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wetfreq')
mu <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wetmean')
nd <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'nv')
ndhr100 <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'count',threshold=100)
ndhr50 <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'count',threshold=50)
q95 <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wq95')
```

@@

```{r, fig.width=6, fig.height=6}
plot(ndhr100,new=FALSE)
grid()
```

And more moderate events

```{r, fig.width=6, fig.height=6}
plot(ndhr50,new=FALSE)
grid()
```

@@

```{r fig.width=6, fig.height=6}
ndhr.cal <- data.frame(y = coredata(subset(subset(ndhr100,is=2),it=c(1949,2004))),
                       X = coredata(subset(eof.es,it=c(1949,2004))))
ds.ndhr <- glm(y ~ X.1 + X.2 + X.3 + X.4 + X.5 + X.6 + X.7, data=ndhr.cal,family="poisson")
plot(1949:2004,ndhr.cal$y,pch=19,ylab=expression(N(X > 100*mm)),xlab='',main=loc(ndhr)[2])
grid()
lines(1949:2004,exp(predict(ds.ndhr)),col='red')
```

Model the number with heavy rainfall (more than 50 mm/day) in Delhi and Bangalore.

```{r fig.width=6, fig.height=6}
## Delhi
ndhr50.1.cal <- data.frame(y = coredata(subset(subset(ndhr50,is=1),it=c(1949,2004))),
                       X = coredata(subset(eof.es,it=c(1949,2004))))
ds.ndhr50.1 <- glm(y ~ X.1 + X.2 + X.3 + X.4 + X.5 + X.6 + X.7, data=ndhr50.1.cal,family="poisson")
plot(1949:2004,ndhr50.1.cal$y,pch=19,ylab=expression(N(X > 100*mm)),xlab='',main=loc(ndhr)[1])
grid()
lines(1949:2004,exp(predict(ds.ndhr50.1)),col='red')

# Bangalore
ndhr50.3.cal <- data.frame(y = coredata(subset(subset(ndhr50,is=3),it=c(1949,2004))),
                       X = coredata(subset(eof.es,it=c(1949,2004))))
ds.ndhr50.3 <- glm(y ~ X.1 + X.2 + X.3 + X.4 + X.5 + X.6 + X.7, data=ndhr50.3.cal,family="poisson")
plot(1949:2004,ndhr50.3.cal$y,pch=19,ylab=expression(N(X > 100*mm)),xlab='',main=loc(ndhr)[3])
grid()
lines(1949:2004,exp(predict(ds.ndhr50.3)),col='red')
```

@@

## Examine the common EOFs for SLP over the Indian sub-continent

```{r}
print('Compare common EOFs')
diag.slp <- biasdiag(slp)
```
